### 2024-10-24 Reports/Final Report/sections/algorithm.tex
final report and codebase added

\begin{algorithm}
\caption{Predicting Responses to Unasked Survey Questions Using Embedding Projections}
\begin{algorithmic}[1]
\State \textbf{Input:} ESS dataset questions, Large Language Model (LLM)
\State \textbf{Output:} Belief embeddings, Predicted answers for omitted questions

\State
\Procedure{QuestionSelectionAndNormalization}{}
    \State Select questions that can be answered on a bipolar/unipolar scale
    \State Exclude questions that cannot be scaled (e.g., "Which party did you vote for?")
    \State Group questions into three categories:
    \begin{itemize}
        \item \textbf{Bipolar scale questions}: Answers range from "Strongly Disagree" to "Strongly Agree"
        \item \textbf{Unipolar scale questions}: Examples include trust levels in politicians
        \item \textbf{Yes/No questions}: Questions with binary answers
    \end{itemize}
    \State Normalize responses:
    \begin{itemize}
        \item Bipolar scale $\rightarrow$ [-1, 1]
        \item Unipolar scale $\rightarrow$ [0, 1]
        \item Yes/No questions: Normalize similar to bipolar scale
    \end{itemize}
\EndProcedure

\State

\Procedure{StatementGeneration}{}
    \For{each question and its response}
        \State \textbf{Use LLM} to transform question-response pairs into statements
        \State \Comment{For example: \textit{"How able do you think you are to take an active role}}
        \State \Comment{\textit{in a group involved with political issues?"}}

        \If{response is \textbf{"missing"} (e.g., "Refusal", "Don't know")}
            \State Generate a statement indicating that the participant did not answer
            \State \Comment{For example: \textit{"I prefer not to answer about my ability to take an}}
            \State \Comment{\textit{active role in a group involved with political issues."}}
        \ElsIf{response is an \textbf{extreme value} (e.g., "Strongly Agree", "Completely able")}
            \State Generate a positive (pivot) statement
            \State \Comment{For example: \textit{"I believe I am completely able to take an active role}}
            \State \Comment{\textit{in a group involved with political issues."}}
        \EndIf
    \EndFor
\EndProcedure

\State

\Procedure{BeliefEmbeddingDerivation}{}
    \For{each participant}
        \State Use LLM to generate semantic embeddings for the statements
        \State Apply normalized coefficients (answers from the first phase) to the embeddings
        \State Aggregate the weighted embeddings to create an individual belief embedding
    \EndFor
\EndProcedure

\State

\Procedure{PredictUnanskedQuestions}{}
    \State Exclude a subset of answered questions from embedding calculation
    \State Use LLM to generate pivot statements of the omitted questions
    \State Use LLM to generate embeddings of the pivot statements
    \State Calculate the projection of each participant's belief embedding onto the target question's vector
    \State Estimate responses based on the distribution of projections
\EndProcedure

\State

\end{algorithmic}
\end{algorithm}