### 2024-11-09 Reports/Final Report/sections/limitations.tex

\section*{Limitations}
Although having a dataset like the ESS provided us with gold annotation for evaluating our predictions at the start of the project, several challenges emerged. Since we had to manually determine whether the questions were unipolar or bipolar, we were limited to conducting experiments only with variables related to politics. It was not enough to decide whether questions were unipolar or bipolar; we also had to exclude questions with nominal response values, such as 'Which party did you vote for?' Additionally, we encountered some difficulties when using open-source language models with the help of the Hugging Face library. While large language models took a long time for inference, smaller models produced inconsistent and irrelevant answers. The variability in the number of response options also made it difficult to approach the problem as a type of classification task by simply adding a standard dense layer when generating predictions.